\newpage
\thispagestyle{plain}

\begin{center}
\fontsize{15}{22.5}\selectfont\bfseries
ACKNOWLEDGEMENTS
\end{center}

\vspace{36pt}

\fontsize{12}{18}\selectfont
I would like to express my sincere gratitude to my thesis advisor, Asst.Prof.Dr. Sarawan Wongsa, for her guidance, support, and expertise throughout this research project. Her insights and encouragement have been invaluable in shaping this work. I also would like to thank Prof. Wuttipong Kumwilaisak for his initiative and thoughtful guidance during the early research phases of this project, and for encouraging innovative methodological approaches.

I am grateful to the Department of Control and Instrumentation Engineering at King Mongkut's University of Technology Thonburi for providing the resources and support necessary for this research. The collaborative environment and academic excellence fostered by the faculty have greatly contributed to my academic journey.

The authors extend heartfelt gratitude to Prof. Adam Brandt, Precourt Institute for Energy, for providing the GasVid dataset, which was instrumental to this research. Special acknowledgment goes to the researchers who developed the GasVid dataset and the original VideoGasNet framework, particularly Jingfan Wang and colleagues, whose foundational work made this research possible.

I also thank Mrs. Phyu Phyu Aung, General Manager at No. 1 Refinery, Thanlyin Myanmar Petrochemical Enterprise, for her expertise on natural gas properties and industrial applications, offering valuable practical insights that enhanced the real-world applicability of this research.

Finally, I would like to thank my family and friends for their unwavering support and encouragement throughout my studies. Their patience and understanding during the demanding periods of research and writing have been a source of strength and motivation.